### 2016-04-04 writing/latex/minimal-document.tex

\documentclass[projectreport,online,forcegraphics,miktex,draft]{ruthesis}
\usepackage[backend=biber,bibencoding=utf8,style=ieee]{biblatex}
\addbibresource{references.bib}
\InputIfExists{custom.tex}
\setTitle{My Title}
\setDocumentType{Final Report}
\author{Firstname1~Lastname1, Firstname1~Lastname2}
\setCourse{T-411-MECH}
\setSchool{\MLSSE{}}
\whensigned{17}{\MLdec}{2015}
\finalifforcegraphics{hyperref} %hyperlinks even in draft mode
\usepackage[hidelinks]{hyperref} 
\begin{document}
\frontmatter{}
\frontcover{}
\enableindents{}
\starttables{} % setup formatting
\tableofcontents{}
\thesislistoffigures{}
\thesislistoftables{}
\starttext{}
\listoffixmes{}
\chapter{Some Chapter}
\section{Some section}
\subsection{Some Subsection}
Some text goes here with a citation~\cite{adams84fish}.
\printbibliography{}
\appendix{}
\input{code} % as an example, perhaps some of your code
\clearforchapter{}
\printindex{}
\backcover{}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: