### 2016-04-22 writing/latex/abstracteng.tex
Add abstract starting

A permutation is an arrangement of \(n\) objects.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. In this thesis, the author provides sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. The author then goes on to completely
classify Wilf-equivalence amongst mesh patterns of length 2 when we avoid the
classical pattern 231.