### 2016-04-30 writing/latex/abstracteng.tex
tidy up a bit

A permutation is an arrangement of \(n\) objects.
Avoidance classes of classical permutation patterns
capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. In this thesis, the author provides sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, the author completely
Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.
% Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
% of sets of the form \(231,m_1\) and \(231,m_2\)