### 2016-04-30 writing/latex/acknowledgements.tex

\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}\sourceatright{Douglas Adams\cite{adams84fish}}
\vspace{\baselineskip}

I would like to thank my advisor, Henning Ulfarsson, for his time, support,
guidance and introducing me to a research topic that I greatly enjoy.

I would like to thank Michael Albert and Anders Claesson for the discussions that
prompted the topic of this thesis, as well as their valuable input and suggestions.

I would like to thank Christian Bean for proofreading my thesis, providing useful
feedback, and for the time spent staring at a whiteboard while I tried to check
my mathematical reasoning.