### 2016-05-04 writing/latex/confut.tex
edit r3

\chapter{Conclusions and Future work}
\nextvar
From \ChapterRef{chap:coincs} it can be seen that automatically classifying
coincidences of mesh patterns is a difficult task, establishing rules for longer
dominating patterns requires many more cases to be taken. It would however
be interesting to consider the application of the third Dominating Rule to mesh
patterns without any dominating pattern in order to try to capture some of the
coincidences described in\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15}
 and \textcite{DBLP:journals/corr/ClaessonTU14}.
It is not possible to apply the first and second Dominating Rules to the pattern
itself, since when applying the rules we consider containers of the pattern inside
the avoiders of the dominating pattern. For example if we were to attempt to apply
the first rule to the pattern \(\perm{1,2}\) then we would have to consider
containers of \(\perm{1,2}\) inside \(\av{\perm{1,2}}\), and obviously this can
never occur.

It is also possible to take sets of mesh patterns instead of a single mesh pattern
when considering dominating rules, and expressing coincidence between these sets.
Doing this may give nice results. Coincidences between sets with
multiple dominating patterns can also be considered, as this provides even more
power to the rules discussed, these methods may be useful for classifying coincidence
in large sets of pattern. It is also possible to apply these rules to sets where the
dominating pattern is a mesh pattern.

It would be interesting to consider a systematic explanation of Wilf-equivalences
amongst classes where \(\perm{3,2,1}\) is the dominating pattern using the
construction presented in \cite[Sec.~12]{2015arXiv151203226B}, in order to directly
reach enumeration and hopefully establish some of the non-trivial Wilf-equivalences
between classes with different dominating patterns. For example, it is possible to show that
the sets \(\scriptvar = \av{\left\{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}\right\}}\)\nextvar and
\(\scriptvar = \av{\left\{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}\right\}}\), are Wilf-equivalent.
This can be seen by considering an occurrence of the mesh pattern,
\(p=\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1}\) in a permutation
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \draw (0.5,1.5) node {\(b_1\)};
        \draw (0.5,2.5) node {\( b_2\)};
        \draw (2.5,2.5) node {\( b_3\)};
        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
    \end{tikzpicture}
    }
\end{equation*}

If we are inside \(\av{\perm{2,3,1}}\) then any points in the region corresponding
to the box \(b_1\) must be to the right of the points in the region corresponding to
box \(b_2\), and must also form a decreasing subsequence by \LemmaRef{lem:incdec}.
Furthermore, the points in the regions corresponding to \(b_2\) and
\(b_3\) must form an avoider of \(\perm{2,3,1}\) with the indicated point.
Therefore the containers of \(p\) in \(\av{\perm{2,3,1}}\) have generating function
\(\lastvar(x) = \frac{x(C(x)-1)}{1-x}\).
Now, if we are \(\av{\perm{3,2,1}}\) then any points in the region corresponding
to the box \(b_1\) must be to the left of the points in the region corresponding to
box \(b_2\), and must also form a increasing subsequence by \LemmaRef{lem:incdec}.
Furthermore, the points in the regions corresponding to \(b_2\) and
\(b_3\) must form an avoider of \(\perm{3,2,1}\) with the indicated point.
Hence, the containers of \(p\) in \(\av{\perm{3,2,1}}\) have generating function
\(\var(x) = \frac{x(C(x)-1)}{1-x}\) and both of these classes have the same
enumeration.