### 2016-05-23 writing/latex/abstracteng.tex
Henning pres & IS abstract

A permutation is an arrangement of \(n\) objects.
Sets of permutations classified by the avoidance of classical permutation
patterns capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. We provide sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases, are used to completely
classify coincidence amongst families containing a mesh pattern of length \(2\)
and a classical pattern of length \(3\).

Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, we completely
Wilf-classify mesh patterns of length \(2\) when avoiding the classical pattern \(231\).
Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
of sets of the form \(231,m_1\) and \(321,m_2\), as well as discussing possible
future work that could be derived from this work.