### 2016-10-24 writing/paper/abstract.tex
Add author and abstract to paper.
Add latexmk files to ignore

\begin{abstract}
    Two mesh patterns are coincident if they are avoided by the same set of permutations, and are
    Wilf-equivalent if they have the same number of avoiders at each length. We provide
    sufficient conditions for coincidence among mesh patterns, whilst also avoiding a longer
    classical pattern. Using theses conditions we completely classify coincidences between families
    containing a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
    completely Wilf-classify equivalences of mesh patterns of length 2 whilst also avoiding the
    classical pattern 231.
\end{abstract}