### 2017-04-05 writing/paper/sections/abstract.tex
Comments from Henning up to section 4

\begin{abstract} Two mesh patterns are coincident if they are avoided by the
same set of permutations, and are Wilf-equivalent if they have the same number
of avoiders of each length. We provide sufficient conditions for coincidence
among mesh patterns, whilst also avoiding a longer classical pattern. Using
these conditions we completely classify coincidences between families containing
a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
completely Wilf-classify equivalences of mesh patterns of length 2 whilst also
avoiding the classical pattern 231.

    \noindent \\ \emph{Keywords: } permutation, pattern, mesh pattern, pattern
    coincidence \end{abstract}