### 2017-04-20 writing/paper/sections/abstract.tex
Final comments from Henning

\begin{abstract}
    Two mesh patterns are coincident if they are avoided by the same set of
    permutations, and are Wilf-equivalent if they have the same number of
    avoiders of each length. We provide sufficient conditions for coincidence of
    mesh patterns, when only permutations also avoiding a longer classical
    pattern are considered. Using these conditions we completely classify
    coincidences between families containing a mesh pattern of length 2 and a
    classical pattern of length 3. Furthermore, we completely Wilf-classify mesh
    patterns of length 2 inside the class of 231-avoiding permutations.

    \noindent \\ \emph{Keywords: } permutation, pattern, mesh pattern, pattern
    coincidence \end{abstract}