### 2017-10-14 writing/paper/sections/coincidences.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

Coincidences among small mesh patterns have been considered by
\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the
Simultaneous Shading Lemma, a closure result and one special case to fully
classify coincidences among mesh patterns of length 2.

Recall that two patterns \(\lambda\) and \(\gamma\) are considered
\emph{coincident} if the set of permutations that avoid \(\lambda\) is the same
as the set of permutations that avoid \(\gamma\), \ie, \(\av{\lambda} =
\av{\gamma}\). This is equivalent to \(\lambda\) and \(\gamma\) being contained
in the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a fixed
classical pattern of length 3 and \(p\) is a mesh pattern of length 2 that
varies. The classical pattern \(\pi\) will be called the \emph{dominating
pattern}. Given such a dominating pattern \(\pi\) we will write \(p_1
\coinc{\pi} p_2\) if for the mesh patterns, \(p_1\) and \(p_2\),  the sets
\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal. If this is the case we say
the two mesh patterns are \emph{coincidenc under} \(\pi\).

Our first step is to calculate whether \(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are
equal up to permutations of length 11 computationally, if they are we write
\(p_1 \coinc[comp]{\pi} p_2\). The equivalence relation \(\coinc{\pi}\) is a
refinement of the equivalence relation \(\coinc[comp]{\pi}\), and therefore the
\(\coinc{\pi}\)-equivalence classes form partitions of the
\(\coinc[comp]{\pi}\)-equivalence classes.

We then prove three propositions called ``Dominating Pattern Rules'', if the
First Dominating Pattern Rule shows that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(1)]{\pi} p_2\). If a combination of the First and Second
Dominating Pattern Rules show that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(2)]{\pi} p_2\); similarly we write \(p_1 \coinc[(3)]{\pi} p_2\) if
a combination of all three rules shows coincidence. These three equivalence
relations are successive refinements of \(\coinc{\pi}\). For future reference
we record the following:

\begin{note}
\label{note:main}
    If we find that \(\coinc[(i)]{\pi}\), for \(i = 1, 2\), or \(3\), equals
    \(\coinc[comp]{\pi}\) then the equivalence relations \(\coinc[(i)]{\pi}\),
    \(\coinc{\pi}\), and \(\coinc[comp]{\pi}\) are all equal.
\end{note}

In order to describe the rules it is useful to have a notion for inserting
points, ascents, and descents into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
\(\boks{i,j}\notin R\). We define a mesh pattern \(p^\boks{i,j} =
(\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a point is
\emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new
underlying classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases}
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:
\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
\(\boks{i,j}\notin R\) and \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as
defined in \DefinitionRef{def:ap}. We define the following four modifications of
\(p^\boks{i,j}\).
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or
bottommost point in \(\boks{i,j}\). We collect the resulting mesh patterns in a
set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
p^\boksleft{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

\begin{figure}
\begin{equation*}
p = \pattern{}{1,2}{0/1,2/2} \quad
p^\boks{2,1}= \pattern{}{1,3,2}{0/1,0/2,2/3,3/3}
\quad
p^\boksup{2,1}
= \pattern{}{1,3,2}{0/1,0/2,2/2,2/3,3/2,3/3}
\end{equation*}
\caption{The result of inserting a point into \(p=(12,\{\boks{0,1},\boks{2,2}\})\)}
\label{fig:addp}
\end{figure}

\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
\(\boks{i,j}\notin R\). We define a mesh pattern \(p^{\boks{i,j}_a} =
(\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\) as the pattern
where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in
\(G(p)\). Formally the new underlying classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+t & \text{if } k = i+t,t\in\{1,2\}\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+2 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-2) & \text{if } \tau(k)\le j \text{ and }k> i+2\\
    \tau(k-2)+2 & \text{if } \tau(k)> j \text{ and }k> i+2\\
\end{cases}
\end{equation*}
The ordering of the top branch determines whether an ascent(or descent) is added.
The mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-2}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-2,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-2,\ell-2}\in R\}\cup\\
&\{\boks{i+1,j},\boks{i+1,j+1},\boks{i+1,j+2}\}
\end{aligned}
\end{equation*}
\end{definition}
An example of adding an ascent to a mesh pattern can be seen in \FigureRef{fig:adda}.
\begin{figure}
\begin{equation*}
p = \pattern{}{2,3,1}{0/0,1/0,1/3,
                      3/0,3/1,3/3} \quad
p^{\boks{1,1}_a}= \pattern{}{4,2,3,5,1}{0/0,
                        1/0,1/5,
                        2/0,2/1,2/2,2/3,2/5,
                        3/0,3/5,
                        5/0,5/1,5/2,5/3,5/5}
\end{equation*}
\caption{The result of inserting an ascent into
\(p=(231,\{\boks{0,0},\boks{1,0},\boks{1,3},\boks{3,0},\boks{3,1},\boks{3,3}\})\)}
\label{fig:adda}
\end{figure}

We now attempt to fully classify coincidences in families characterised by
avoidance of a classical pattern of length \(3\) and a mesh pattern of length
\(2\), that is finding and explaining all coincidences between mesh patterns
\(m\) and \(m^\prime\), \(m\coinc{\pi} m^\prime\), where \(\pi\) is a classical
pattern of length \(3\).

It can be easily seen that in order to classify coincidences one need only
consider coincidences within the family of mesh patterns with the same
underlying classical pattern, this is due to the fact that
\(\perm{2,1}\in\av{\mperm{1,2}{R}}\) and
\(\perm{1,2}\in\av{\{\mperm{2,1}{R}\}}\) for all mesh-sets \(R\).

We know that there are a total of \(512\) mesh-sets for each underlying
classical pattern. By use of the previous results of
\textcite{DBLP:journals/corr/ClaessonTU14} the number of coincidence classes can
be reduced to \(220\).

\subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}
Through experimentation, considering avoidance of permutations up to length
\(11\), we discover that there are \(29\) \(\coinc[comp]{321}\)-equivalence
classes where the underlying classical pattern of the mesh pattern is \(21\).

\begin{proposition}[First Dominating Pattern Rule]
    \label{prop:dom1}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 1\), then \(m_1 \coinc{\pi} m_2 \)
    if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}

This rule can be understood in graphical form. In the pattern in
\FigureRef{fig:rule1} we can gain shading in the boxes \(\boks{0,2}\), \(\boks{2,0}\)
since if there is a point in either of these boxes there would be an occurrence of
the dominating pattern \(\perm{3,2,1}\).

\begin{figure}[htb]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
    \end{tikzpicture}

        \caption{Visual depiction of first dominating pattern rule.}
        \label{fig:rule1}
    \end{center}
\end{figure}

In order to prove the proposition we must first make the following note.

\begin{note}
    \label{not:downcmesh}
    Let \(R_1 \subseteq R_2\). Then any occurrence of \((\tau, R_2)\) in a permutation
    is an occurrence of \((\tau,R_1)\).
\end{note}

\begin{proof}[Proof of \PropositionRef{prop:dom1}]
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} implies that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now we consider a permutation \(\omega \in \av{\pi}\), containing an
    occurrence of \(m_1\). If there is a point in the region corresponding to
    the box \(\boks{a,b}\), then that point, along with the points of the
    occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\). Then
    condition~\eqref{prop:dom1:cont} of the proposition implies an occurrence of
    \(\pi\). Therefore there can be no points in this region, which implies the
    occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

    Taking both directions of the containment we can therefore draw the
    conclusion that \(m_1 \coinc{\pi} m_2\).
\end{proof}

After implementing the First Dominating Pattern Rule we find that there are
\(29\) \(\coinc[(1)]{321}\)-equivalence classes, of mesh patterns where the
underlying classical pattern is \(21\). By \NoteRef{note:main} there are
therefore exactly \(29\) \(\coinc{321}\)-equivalence classes.

\subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
By application of First Dominating Pattern Rule we obtain \(43\)
\(\coinc[(1)]{231}\)-equivalence classes between mesh patterns with \(21\) as an
underlying classical pattern. Experimentation shows that there are \(39
\coinc[comp]{231}\)-equivalence classes between mesh patterns with \(21\) as an
underlying classical pattern, for example the following two patterns are in the
same \(\coinc[comp]{231}\)-equivalence class, but this is not explained by the
First Dominating Pattern Rule.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
\end{equation*}

Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\),
consisting of elements \(x\) and \(y\). If the region corresponding to the box
\(\boks{1,1}\) is empty we have an occurrence of \(m_2\). Otherwise, if there is
any ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since the permutation is in \(\av{\perm{2,3,1}}\) this is not possible.
This box must therefore contain a (non-empty) decreasing subsequence. This gives
rise to the following lemma:

\begin{lemma}
    \label{lem:incdec}
    Let \(m =(\sigma, R)\) be a mesh pattern, where \(\boks{a,b} \notin R\),
    and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
    \(\pi \preceq m^{\boks{a,b}_a}\)\\(\(\pi \preceq m^{\boks{a,b}_d}\)),
    then in any occurrence of \(m\) in a permutation \(\varrho\), the region
    corresponding to the box \(\boks{a,b}\) can only contain an decreasing
    (increasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/1,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
we know that the region corresponding to the box \(\boks{1,1}\) contains a
decreasing subsequence. If we let \(z\) be the topmost point in this decreasing
subsequence, then \(xz\) is an occurrence of \(m_2\). This shows that our two
example patterns are coincident. This example generalises into the following
rule.

\begin{proposition}[Second Dominating Pattern Rule]
    \label{prop:dom2}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 2\), then \(m_1 \coinc{\pi} m_2 \)
    if
\begin{enumerate}
  \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
  \item\label{prop:dom2:cond} Any one of the following four conditions hold
    \begin{enumerate}
    \item\label{prop:dom2:condc} \(\pi \preceq \sigma^{\boks{a,b}_a}\) and
      \begin{enumerate}
        \item \((a+1,b) \in \sigma\) and \(\boks{a+1,b-1}\notin R_1\) and \\
          \(\boks{x,b-1}\in R_1 \implies \boks{x,b} \in R_1 \) (where \(x\neq a,a+1\)) and\\
          \(\boks{a+1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b-1,b\)).
        \item \((a,b+1) \in \sigma\) and \(\boks{a-1,b+1}\notin R_1\) and \\
          \(\boks{x,b+1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a-1,a\)) and\\
          \(\boks{a-1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b,b+1\)).
      \end{enumerate}
    \item \(\pi \preceq \sigma^{\boks{a,b}_d}\) and
      \begin{enumerate}
        \item \((a+1,b+1) \in \sigma\) and \(\boks{a+1,b+1}\notin R_1\) and \\
          \(\boks{x,b+1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a,a+1\)) and\\
          \(\boks{a+1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b,b+1\)).
        \item \((a,b) \in \sigma\) and \(\boks{a-1,b-1}\notin R_1\) and \\
          \(\boks{x,b-1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a-1,a\)) and\\
          \(\boks{a-1,y}\in R_1 \implies \boks{a,y} \in R_1\)  (where \(y\neq b-1,b\)).
      \end{enumerate}
    \end{enumerate}
\end{enumerate}
\end{proposition}
\begin{proof}
    By \NoteRef{not:downcmesh} we only need to show that an occurrence of
    \(m_1\) implies an occurrence of \(m_2\). We consider taking the first
    branch of every choice in condition~\eqref{prop:dom2:cond}. Consider a
    permutation \(\omega \in \av{\pi}\). Suppose \(\omega\) contains \(m_1\) and
    consider the region corresponding to \(\boks{a,b}\) in \(R_1\). If the
    region is empty, the occurrence of \(m_1\) is trivially an occurrence of
    \(m_2\).

    If the region is non-empty, then by \LemmaRef{lem:incdec} and
    condition~\eqref{prop:dom2:condc} of the proposition it must contain a
    decreasing subsequence. We can choose the topmost point in the region to
    replace the corresponding point in the mesh pattern and the points from the
    subsequence are now in the box southeast of the point.

    The conditions on the mesh ensure that no elements of the permutation that
    were inside a region corresponding to an unshaded box in the occurrence of
    \(m_1\) would be in a region corresponding to a shaded box in an occurrence
    of \(m_2\).

    Hence there are no points in the region corresponding to the box
    \(\boks{a,b}\) in the mesh pattern, and therefore we can shade this region.
    This implies that every occurrence of \(m_1\) in \(\av{\pi}\) can be
    modified into an occurrence of \(m_2\) so \(\av{\{\pi,m_2\}} \subseteq
    \av{\{\pi,m_1\}}\).

    Similar arguments cover the remainder of the branches.
\end{proof}

This proposition essentially states that we slide all of the points in the box
we desire to shade diagonally, and chose the topmost/bottommost point to replace
the original point in the mesh pattern. \FigureRef{fig:d2} show how the cases
apply to a general container of \(m_1\) to transform it into a container of
\(m_2\), the circled points are the same points in the permutation.
\begin{figure}
\begin{tikzpicture}
	\modpattern[3]{1,2}{2,1}{0/0,0/1,0/2,1/0,2/1}
	\filldraw (1.33,1.66) circle (1.5pt);
	\filldraw (1.66,1.33) circle (1.5pt);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.1,0.9) rectangle (2.5,0.0);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.5,3) rectangle (2.9 ,2.1);
\end{tikzpicture}
\raisebox{8ex}{\(\coinc{}\)}
\begin{tikzpicture}
	\modpattern[3]{1}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
	\filldraw (2.15,0.85) circle (1.5pt);
	\filldraw (2.3,0.7) circle (1.5pt);
	\draw (2.3,0.7) circle (3pt);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.4,0.6) rectangle (2.65,0.0);
	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.65,3) rectangle (2.9 ,2.1);
\end{tikzpicture}
\caption{A depiction of the second dominating pattern rule using our example patterns.}
\label{fig:d2}
\end{figure}

After implementing the Second Dominating Pattern Rule we find that there are
\(39\) \(\coinc[(2)]{231}\)-equivalence classes, of mesh patterns where the
underlying classical pattern is \(21\). By \NoteRef{note:main} there are
therefore exactly \(39\) \(\coinc{231}\)-equivalence classes.

\subsection{Coincidence classes of Av(\{231, (12, \textit{R})\}).}
When considering the coincidence classes of
\(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\) we first apply the two Dominating
Pattern rules previously established. Starting from \(220\) classes, we result
in \(85 \coinc[(1)]{231}\)-equivalence classes reducing to \(59
\coinc[(2)]{231}\)-equivalence classes, considering mesh patterns with \(12\) as
the underlying classical patterns. However there are fewer
\(\coinc[comp]{231}\)-equivalence classes.

For example the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are \(\coinc[comp]{231}\)-equivalent.

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now
look at the case when this region is not empty, and consider choosing the
rightmost point in the region. This gives us an occurrence of the following mesh
pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of First Dominating Pattern Rule we obtain the following mesh
pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
The highlighted points are an occurrence of the mesh pattern \(m_2\). This gives
rise to the following rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
    \(m_1 \coinc{\pi} m_2\) if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(m_2\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
as before. Now consider a permutation \(\omega \in \av{\pi}\) that contains an
occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\) is
empty then we have an occurrence of \(m_2\). By
condition~\ref{prop:dom3:condocc} of the proposition there exists a pattern
\(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some pattern
\(r'\) coincident with \(r\). If the region is non-empty then there exists an
occurrence of \(r\) in \(\omega\). Therefore, there is an occurrence of \(r'\) in
\(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the two patterns are
coincident.
\end{proof}

After implementing the Third Dominating Pattern Rule we find that there are
\(56\) \(\coinc[(3)]{231}\)-equivalence classes, of mesh patterns where the
underlying classical pattern is \(12\). By \NoteRef{note:main} there are
therefore exactly \(56\) \(\coinc{231}\)-equivalence classes, since we have
observed \(56\) \(\coinc[comp]{231}\)-equivalence classes through
experimentation.

\subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
When considering coincidences of mesh patterns with underlying classical pattern
\(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules gives no coincidences. Through experimentation we discover
that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes that are
unexplained. Since the number of coincidences is so small we will reason for
these coincidences without attempting to generalise into concrete rules.

Intuitively, it is easy to see why our previous rules have no power here. It is
impossible to add a single point to a mesh pattern \((12, R)\) and create an
occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have a position
where addition of an ascent, or descent, provides extra shading power.

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \et m_2 \pattern{}{1,2}{0/0,
    0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are \(\coinc[comp]{321}\)-equivalent. (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

Consider the region corresponding to the box \(\boks{0,1}\) in any occurrence of
\(m_1\) in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing
subsequence. If the region is empty then we have an occurrence of \(m_2\). If
there is only one point in the region we can choose this point to replace the
\(1\) in the mesh pattern to get the required shading. If there is more than one
point then choosing the two leftmost points gives us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Here the two highlighted points are the original two points. The other two
points are an occurrence of the pattern we \(m_2\), and hence the two patterns
are coincident. It is also possible to calculate this coincidence by an
extension of the Third Dominating rule, where we allow a sequence of point
addition operations, this is discussed further in the future work section.
Because of symmetries we have dealt with \(4\) out of the \(7\)
\(\coinc[comp]{321}\)-equivalence classes.

Now consider the patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
\end{equation*}
which are \(\coinc[comp]{321}\)-equivalent.

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in the region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, giving the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty
        then we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern as any
        points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an
        occurrence of the dominating pattern \(\perm{3,2,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore we have
explained all 7 of the coincidences in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\)
and there are 213 \(\coinc{321}\)-equivalence classes.