### 2017-10-14 writing/paper/sections/introduction.tex

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. The original concept---a subsequence of symbols
having a particular relative order, now known as classical patterns---has been
expanded to a variety of definitions. \textcite{babstein2000} considered
\emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed}
patterns) where two adjacent entries in the pattern can be required to be
adjacent in the permutation. \textcite{MR2652101} look at classes of patterns
where entries can also be required to be consecutive in value, these are called
\emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by
\textcite{MR2264071} to establish necessary conditions for a Schubert variety to
be Gorenstein. These definitions are subsumed under the definition of \emph{mesh
patterns}, introduced by \textcite{journals/combinatorics/BrandenC11} to capture
explicit expansions for certain permutation statistics.

When considering permutation patterns some of the main questions posed relate to
how and when a pattern is avoided by, or contained in, an arbitrary set of
permutations. Two patterns \(\pi\) and \(\sigma\) are \emph{Wilf-equivalent} if
the number of permutations that avoid \(\pi\) of length \(n\) is equal to the
number of permutations that avoid \(\sigma\) of length \(n\). A stronger
equivalence condition is that of \emph{coincidence}, where the set of
permutations avoiding \(\pi\) is exactly equal to the set of permutations
avoiding \(\sigma\). Avoiding pairs of patterns of the same length with certain
properties has been studied, \textcite{MR2178749} considered avoiding a pair of
vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding a
vincular and a covincular pattern simultaneously in order to achieve several
counting results. However, little work has been done on avoiding a mesh pattern
and a classical pattern simultaneously.

In this work we aim to establish some ground in this field by computing
coincidences and Wilf-classes and calculating some of the enumerations of
avoiders of a mesh pattern of length 2 and a classical pattern of length 3. We
begin by establishing coincidences between mesh patterns of length 2 while
avoiding a dominating pattern by computational methods, which are then used to
establish three ``Dominating Pattern Rules'' as well as some special cases that
can be used to calculate coincidences.

 We then use these coincidence classes to calculate Wilf-equivalence classes
 showing some of the methods used.