### 2017-10-14 writing/paper/sections/meshpatterns.tex

A \emph{permutation} is a bijection from the set \(\nrange{n} =
\setrange{1}{n}\) to itself. The set of all such bijections is denoted
\(\mathfrak{S}_n\) and has \(n!\) elements. We can denote an individual
permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
entries of the permutation in order, therefore \(\pi =
\perm{\pi(1),\pi(2),\dotsm,\pi(n)}\). The set \(\mathfrak{S}_0\) has exactly one
element, the empty permutation \(\varepsilon\).

\begin{definition}
    Two strings of integers \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

The definition of order isomorphism allows us to give the meaning of containment for classical
permutation patterns.
\begin{definition}
    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation
    \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted} \(\sigma \preceq \pi\)) if there
    is some sequence \( i_1,i_2,\dotsc,i_k\) such that \(1\le
    i_1<i_2<\dotsm<i_k\le n\) and the sequence
    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If this is the case the sequence
    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of
    \(\sigma\) in \(\pi\). If \(\pi\) does not contain \(\sigma\), we say that
    \(\pi\) \emph{avoids} \(\sigma\). In this context \(\sigma\) is called a
    (\emph{classical}) \emph{permutation pattern}.
\end{definition}

\begin{example}
\label{ex:contexmpl}
The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern \(\sigma =
\perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order
isomorphic to \(\perm{231}\). The permutation also contains the occurrence
\(241\) of the same pattern. The permutation \(\perm{2,4,1,5,3}\) avoids the
pattern \(\perm{3,2,1}\).
\end{example}

We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\)
as \(\Av_n(\sigma)\) and let
\(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).

We can display a permutation graphically in a \emph{plot}, where we display the
points \(G(\pi) = \{(i,\pi(i))\mid i \in \nrange{n} \}\) in a Cartesian
coordinate system. The plots of the permutations \(\pi = \perm{2,4,1,5,3}\) and
\(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
\FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in
\ExampleRef{ex:contexmpl}.

The boxes in the plot of a permutation are denoted by \(\boks{i,j}\), where the
point \((i,j)\) is the lower left corner of the box.

\begin{figure}[htb]
    \begin{center}
    \raisebox{6ex}{\(G(\pi)=\)}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,4,1,5,3}{}
    \end{tikzpicture}
    \raisebox{6ex}{\(\quad{}G(\sigma)=\)}
    \raisebox{0.35cm}{
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,3,1}{}
    \end{tikzpicture}}

        \caption{The plots of the permutations \(\pi\) and \(\sigma\).}
        \label{fig:plots}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \(\pattern{2,4,5}{2,4,1,5,3}{}\)
        \caption{The occurrence of 231 in 24153 corresponding to 453.}
        \label{fig:containment}
    \end{center}
\end{figure}

\begin{definition}
A \emph{mesh pattern} is a pair
\begin{equation*}
    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq [0,k]\times [0,k].
\end{equation*}
\end{definition}

Formally defined by \textcite{journals/combinatorics/BrandenC11}, an
\emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are
order-preserving injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\)
satisfying the following two conditions.
Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
  \begin{enumerate}[i.]
    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\).
  \end{enumerate}\saveenum
  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
   the second condition is
  \begin{enumerate}[i.]\resetenum
  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\).
\end{enumerate}
We call \(R_{ij}\) the \emph{region corresponding to} \(\boks{i,j}\).

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{\boks{0,1},\boks{0,2},\boks{1,0},\boks{1,1},\boks{2,1},\boks{2,2}\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }\) is contained in
    \(\pi = \perm{3,4,2,1,5}\).
    Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
    an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
    are not contained in the regions corresponding to the shaded boxes in \(p\).
    \begin{equation*}
        \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                                   1/0,1/1,1/2,
                                   2/0,2/1,2/2,
                                   3/2,3/3,3/4,
                                   4/2,4/3,4/4}
    \end{equation*}
    The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\).
\end{example}

We denote the avoidance sets for mesh patterns in the same way as for classical
patterns. Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the
\emph{underlying classical pattern} of \(p\).

We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as
above, with the additional condition that if \(\boks{i,j}\in R\text{ then }
R_{ij}\) is contained in the mesh set of \(q\). More formally:

\begin{definition}
A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as
a \emph{subpattern} if \(\kappa\) contains \(p\) and
\(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
\end{definition}

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{\boks{0,1},\boks{0,2},\boks{1,0},\boks{2,2}\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 2/2 }\) is contained in the pattern
    \(q = \textpattern{}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                               1/0,1/1,1/2,1/3,1/4,
                               2/1,2/2,2/3,2/4,
                               3/2,3/4,
                               4/2,4/3,4/4} \) as a subpattern.
    The highlighted points form an occurrence of \(p\) in \(q\)
    \begin{equation*}
        \pattern{1,2,5}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                                   1/0,1/1,1/2,1/3,1/4,
                                   2/1,2/2,2/3,2/4,
                                   3/2,3/4,
                                   4/2,4/3,4/4}
    \end{equation*}
    The permutation \(\perm{4,2,3,1,5}\) also contains \(p\) in the usual sense.
\end{example}